---
title: 'Statistical Inference: ToothGrowth'
author: "Francisco Marco-Serrano"
output: pdf_document
---

#Summary
According to R Help, the ToothGrowth dataset contains the information on the length of teeth in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid). We analyse the differences in length growth by dose level and delivery method.

#Data Exploratory Analysis
```{r}
library(datasets)
library(lattice)

data.fr <- ToothGrowth

# Basic information on dimension of the data frame and the variables
dim(data.fr)
str(data.fr)

# including summary
summary(data.fr)

# mean by dose and supplement
mag <- aggregate(len ~ ., data = data.fr, mean)
xtabs(len ~ ., data = mag)

# var by dose and supplement
vag <- aggregate(len ~ ., data = data.fr, var)
xtabs(len ~ ., data = vag)

# histogram of tooth growth
histogram( ~ data.fr$len, data = data.fr,
           main = "Histogram", xlab = "Tooth length")

# histogram per dose and type of supplement
data.fr$dose <- as.factor(data.fr$dose)
histogram( ~ data.fr$len | data.fr$dose+data.fr$supp, data = data.fr,
           main = "Histogram of tooth length by Supplement and Dose",
           xlab = "Tooth length")

# and boxplot per dose and type of supplement
boxplot(data.fr$len ~ data.fr$dose + data.fr$supp, horizontal = TRUE)

```

The basic exploratory analysis suggests there are differences in the tooth length depending on the dose and type of supplement.


#Confidence Intervals

We test the difference in the means of tooth length between the group that received one type of supplement and the other group.

```{r}

# Independent groups T-test with unequal variance
# for assessing differences from supplement
t.test(data.fr$len~data.fr$supp, var.equal = FALSE)

# For assessing differences from dose, an ANOVA model would be better,
# however we are instructed to use T-tests:

# 0.5 vs rest
data.fr$d05 <- 0
data.fr$d05[which(data.fr$dose == "0.5")] <- 1
aggregate(len ~ d05, data = data.fr, var)
t.test(data.fr$len~data.fr$d05, var.equal = FALSE)

# 1.0 vs rest
data.fr$d10 <- 0
data.fr$d10[which(data.fr$dose == "1")] <- 1
aggregate(len ~ d10, data = data.fr, var)
t.test(data.fr$len~data.fr$d10, var.equal = FALSE)

# 2.0 vs rest
data.fr$d20 <- 0
data.fr$d20[which(data.fr$dose == "2")] <- 1
aggregate(len ~ d20, data = data.fr, var)
t.test(data.fr$len~data.fr$d20, var.equal = FALSE)

```

#Conclusions
* With a 95% confidence, there is no conclusive evidence the type of supplement explains the difference in tooth growth.
* From the p-value, when considering the type of supplement, differences in tooth growth could be considered at significance levels above 6.1%.
* From the p-value, when considering the dose, differences in tooth growth could be considered from 0.5mg vs the rest of doses.
* There is no significant difference in tooth growth when considering a 1mg dose vs the rest of doses.
* From the p-value, when considering the dose, differences in tooth growth could be considered from 2.0mg vs the rest of doses.
