---
title: 'Most Severe Weather Events in the US: Health and Economic Issues'
author: "Francisco Marco-Serrano"
output: html_document
---

#Synopsis

Since 1996, year from which we have full information on the current 48 events NOAA is recording major storms and weather events information, there have been 653530 of such events in the US.
On average, the type of event most harmful with respect to population health is *Heat Wave*, with a combined average of fatalities and injuries of 70 per recorded event. As per the economic consequences, *HURRICANE/TYPHOON* is the event generating most losses on average, with a 817.20 M$.

#Data Processing

From the [U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database](www.ncdc.noaa.gov/stormevents) we obtained data on major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage since 1950 until November 2011.

We first read the data from the bz2 compressed file. The data is a CSV file containing 902297 rows and 37 variables.

```{r cache = TRUE}
storm.df = read.csv("repdata-data-StormData.csv.bz2")

dim(storm.df)  # how big the data frame is
str(storm.df)  # which is the structure
head(storm.df) # show first few rows
 
```

We make sure the date is formated adequately, and create a new variable (YEAR). Keep only the variables for YEAR, STATE, Event Type, Population Health (FATALITIES plus INJURIES), and Economic Consequences (PROPDMG & PROPDMGEXP for property damages plus CROPDMG & CROPDMGEXP for crop damages).

Only data from 1996 to present is used in the analysis since before not all events were recorded. These are a total of 653530 events.


```{r}
storm.df$BGN_DATE = as.Date(storm.df$BGN_DATE, format = "%m/%d/%Y %H:%M:%S")
storm.df$YEAR = as.integer(format(storm.df$BGN_DATE, "%Y")) 
 
# subset, to keep only the variables we are interested in
keeps = c("YEAR", "STATE", "EVTYPE", "FATALITIES", "INJURIES", "PROPDMG", "PROPDMGEXP", "CROPDMG", "CROPDMGEXP")
storm.df = storm.df[storm.df$YEAR > 1995 , keeps]

dim(storm.df)

```

We clean the event types (NOOA states only 48 types, while the database reflects 985).

```{r}
# Trim spaces
library(stringr)
storm.df$EVTYPE = str_trim(storm.df$EVTYPE)

# Normalise some events...
pattern = "(THUNDERSTORM|TSTM)"
events.index = grep(pattern, storm.df$EVTYPE)
storm.df$EVTYPE[events.index] = "THUNDERSTORM"

pattern = "(FLOOD)"
events.index = grep(pattern, storm.df$EVTYPE)
storm.df$EVTYPE[events.index] = "FLOOD"

```

We transform PROPDMG and CROPDMG in USD$ according to PROPDMGEXP and CROPDMGEXP (K for thousands, M for millions, B for billions).

```{r}

# Normalise PROPDMG & CROPDMG according to the EXP variables
# Calculated in millions
pattern = "(K|k)"
exp.index = grep(pattern, storm.df$PROPDMGEXP)
storm.df$PROPDMG[exp.index] = storm.df$PROPDMG[exp.index] * 0.001
exp.index = grep(pattern, storm.df$CROPDMGEXP)
storm.df$CROPDMG[exp.index] = storm.df$CROPDMG[exp.index] * 0.001

pattern = "(M|m)"
exp.index = grep(pattern, storm.df$PROPDMGEXP)
storm.df$PROPDMG[exp.index] = storm.df$PROPDMG[exp.index] * 1
exp.index = grep(pattern, storm.df$CROPDMGEXP)
storm.df$CROPDMG[exp.index] = storm.df$CROPDMG[exp.index] * 1

pattern = "(B|b)"
exp.index = grep(pattern, storm.df$PROPDMGEXP)
storm.df$PROPDMG[exp.index] = storm.df$PROPDMG[exp.index] * 1000
exp.index = grep(pattern, storm.df$CROPDMGEXP)
storm.df$CROPDMG[exp.index] = storm.df$CROPDMG[exp.index] * 1000

```

We group the data adding the four variables of interest per type of event.

```{r}

# mean of the 1996-2011 period per event
storm.mean = aggregate(cbind(FATALITIES, INJURIES, PROPDMG, CROPDMG) ~ EVTYPE, data = storm.df, FUN="mean")

```

#Results

Show Top 10 events with higher average values for each one of the four variables.

```{r}

head(storm.mean[order(-storm.mean$FATALITIES),], 10)
head(storm.mean[order(-storm.mean$INJURIES),], 10)
head(storm.mean[order(-storm.mean$PROPDMG),], 10)
head(storm.mean[order(-storm.mean$CROPDMG),], 10)

```

We generate the 'Population Health' variable adding FATALITIES and INJURIES, and the 'Economic Consequences' variable adding PROPDMG and CROPDMG.

```{r}

storm.mean$POPHEALTH = storm.mean$FATALITIES + storm.mean$INJURIES
storm.mean$ECONCONS = storm.mean$PROPDMG + storm.mean$CROPDMG

```

Show Top 10 events most harmful for population health and with worst economic consequences.

```{r}

storm.top10H = storm.mean[order(-storm.mean$POPHEALTH),][1:10, ]
storm.top10H
barplot(storm.top10H$POPHEALTH,
        main = "Average total fatalities and injuries (Top 10 Events)",
        names.arg = storm.top10H$EVTYPE, las = 2)


storm.top10E = storm.mean[order(-storm.mean$ECONCONS),][1:10, ]
storm.top10E
barplot(storm.top10E$ECONCONS,
        main = "Average economic consequences M.US$ (Top 10 Events)",
        names.arg = storm.top10E$EVTYPE, las = 2)

```
