---
title: "Regression Models Assignment: Motor Trend Case Study"
author: "Francisco Marco-Serrano"
output: pdf_document
---

# Executive Summary
Eleven variables are analised for thirty-two cars in search of differences in consumption for cars with automatic and manual transmission. A difference of 2.94 miles per gallon is found in manual transmission car, a difference significant at a 5% level (95% confidence).

# Dataset and Exploratory Analysis
The analysed dataset is the `mtcars` data from the `datasets` package. It contains information on 11 variables for 32 observations from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models). See variables description in the Appendix.

```{r echo=FALSE}
library(datasets)
```

The variables are correlated in the following manner:

```{r echo = FALSE}
library(corrgram)
corrgram(mtcars, order = TRUE,
         lower.panel = panel.shade,
         upper.panel = panel.pts,
         text.panel = panel.txt)
```

As we can see, all five numerical variables seem to have either an indirect relationship with the miles per gallon (i.e. disp, hp, wt) or a direct one (i.e. drat, qsec).


# Analysis

Different nested models are tested starting with all the variables as confounders. Significance tests and comparison of adjusted coeficient of determination are used to assess the suitability of the model. See Appendix for detailed results.

Finally, to the best fitted model is shown below.

```{r}
fit <- lm(mtcars$mpg ~ mtcars$wt + mtcars$qsec + factor(mtcars$am), mtcars)
summary(fit)
```

The variable `am` is significant at any 5% significance level (p-value = 0.046716, which suggests that at 1% we would not be rejecting the null hypothesis). As the variable is defined for taking unitary values for manual transmision, the positive coefficient points out towards a better `mpg` for the manual transmision cars versus the consumption of the automatic transmision ones. The mean difference is of 2.9358 miles/(US) gallon.

## Plot the residuals and run diagnostics for normality, homocedasticity and multicollinearity.


```{r}
fit.fitted <- fitted(fit)
fit.resid <- resid(fit)
plot(fit.fitted, fit.resid,
     xlab = "fitted values", ylab = "residuals", main = "Residual Plot")
abline(0, 0)
```

Diagnose normality by using the QQ plot: the diagram suggests the error term is normally distributed.

```{r}
qqPlot(fit, main="QQ Plot")
```

Diagnose homocedasticity by using the Breusch-Pagan test: the p-value is 0.1029 so if we consider a significance level of 5%, we could not reject the hypothesis of homocedasticity (at a 10%, we would be in a border line situation).

```{r}
library(lmtest)
bptest(fit)
```

Diagnose multicollinearity using the Variance Inflation Factors (VIF): none of the values exceeds the rule of thumb of 2. So, the presence of multicollinearity is not considered.

```{r}
library(car)
sqrt(vif(fit))
```

# Appendix

## Variables
Variable | Description
---------|------------
mpg      | Miles/(US) gallon
cyl      | Number of cylinders
disp     | Displacement (cu.in.)
hp       | Gross horsepower
drat     | Rear axle ratio
wt       | Weight (lb/1000)
qsec     | 1/4 mile time
vs       | V/S
am       | Transmission (0 = automatic, 1 = manual)
gear     | Number of forward gears
carb     | Number of carburettors


```{r}
# Show data frame structure
str(mtcars)
```


## Model Selection
```{r}
x<-mtcars[c(1:11)]        # numeric variables
fit1 <- lm(x$mpg ~.,x)
summary(fit1)

x<-mtcars[c(1,3,4,6,7,9)] # next best model by reducing non-significant variables
fit2 <- lm(x$mpg ~.,x)
summary(fit2)

x<-mtcars[c(1,6,7,9)]     # simplest model by reducing non-significant variables
fit3 <- lm(x$mpg ~.,x)
summary(fit3)
```

Selected model is `fit3`, which explains mpg by wt, qsec and am. All three variables significant at 5% level, and an adjusted coeficient of determination of 83.36% (unadjusted is 84.97%).