---
title: "Data Science Capstone - Milestone Report"
author: "Francisco Marco-Serrano"
output: html_document
---

#Briefing
The aim of this report is to describe the three files that will be used to build the corpus employed to model a predictive algorithm for Swiftkey. The data is from a corpus called HC Corpora ([www.corpora.heliohost.org]) and it contains samples from news, blogs, and Twitter, in English, German, Russian, and Finnish.

#Data
In this report, the English datasets are explored, and prepared for the modelling.

```{r}
#Load required libraries
library(tm)
library(RWeka)
library(ggplot2)

#Load data
myfile <- "./final/en_US/en_US.blogs.txt"
en_US.blogs <- scan(file=myfile, what="character", sep="\n", quote="")

myfile <- "./final/en_US/en_US.news.txt"
en_US.news <- scan(file=myfile, what="character", sep="\n", quote="")

myfile <- "./final/en_US/en_US.twitter.txt"
en_US.twitter <- scan(file=myfile, what="character", sep="\n", quote="")

rm(myfile) #used to save memory

#Some basic summaries of the three files
##Number of elements/documents (lines)
length(en_US.blogs)
length(en_US.news)
length(en_US.twitter)

##Descriptive stats on the number of characters of the documents (basic data tables)
summary(nchar(en_US.blogs))
summary(nchar(en_US.news))
summary(nchar(en_US.twitter))

##Word count
sum(sapply(strsplit(en_US.blogs, " "), length))
sum(sapply(strsplit(en_US.news, " "), length))
sum(sapply(strsplit(en_US.twitter, " "), length))

```

#Analysis of the documents

````{r}
#Select a sample of 10% of the total documents
blogs <- sample(en_US.blogs, size = ceiling(length(en_US.blogs)/10),
                      replace = FALSE)
news <- sample(en_US.news, size = ceiling(length(en_US.news)/10),
                      replace = FALSE)
twitter <- sample(en_US.twitter, size = ceiling(length(en_US.twitter)/10),
                      replace = FALSE)

rm(en_US.blogs, en_US.news, en_US.twitter)

#Generate corpus (text that will be used to build algorithm)
corp.source <- VectorSource(paste(blogs, news, twitter))
rm(blogs, news, twitter)

corpus <- VCorpus(corp.source, readerControl = list(language = "English"))
rm(corp.source)

#Clean corpus; use getTransformations() to see all available.
##every word to lower case
corpus <- tm_map(corpus, content_transformer(tolower))
##remove numbers
corpus <- tm_map(corpus, removeNumbers)
##strip whitespaces
corpus <- tm_map(corpus, stripWhitespace)
##remove stopwords
corpus <- tm_map(corpus, removeWords, stopwords("english"))

##remove own stopwords
##profanity list from http://www.cs.cmu.edu/~biglou/resources/
url <- "http://www.cs.cmu.edu/~biglou/resources/bad-words.txt"
profanity <- read.csv(url, header=FALSE)
ownStopWords <- as.character(profanity[[1]])
corpus <- tm_map(corpus, removeWords, ownStopWords)

rm(url, profanity, ownStopWords)

#Get the cleaned term matrix
dtm <- DocumentTermMatrix(corpus)
dim(dtm)

#Remove sparse items
dtm2 <- removeSparseTerms(dtm, sparse=0.95)
dim(dtm2)

rm(dtm)

#Show frequent terms
freq <- sort(colSums(as.matrix(dtm2)), decreasing=TRUE)
word.freq <- data.frame(word=names(freq), freq=freq)
barplot(height=word.freq$freq, names.arg=word.freq$word)

```

#Next milestones

Considering the results shown before and the size of the corpus, we are planning the following milestones for developing the predicting algorithm:

*Work with different sample sizes and techniques of the files.
*Clean better the Twitter documents.
*Tokenise the corpus and build N-grams.
*Identify a suitable technique for modelling the English language.
*Deploy algorithm in a Shinny web-app.
