---
title: "Machine Learning Assignment"
author: "Francisco Marco-Serrano"
output: html_document
---

##Synopsis
A machine learning algorithm is devised to predict activity quality from activity monitors. It is described what the out of sample error is and an estimate of the error is given from a cross-validation exercise.

##Data and exploratory analysis

The dataset has 5 classes (sitting-down, standing-up, standing, walking, and sitting) collected on 8 hours of activities of 4 healthy subjects. The authors also established a baseline performance index.

Read more: http://groupware.les.inf.puc-rio.br/har

*Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6.*


```{r}
library(caret)
library(rattle)
library(rpart.plot)

train.df <- read.csv("pml-training.csv", na.strings = "NA")
test.df <- read.csv("pml-testing.csv", na.strings = "NA")

# Take out summary variables
toremove <- c(1:7, 12:36, 50:59, 69:83, 87:112, 125:150)
train.df <- train.df[ , -toremove]
test.df <- test.df[ , -toremove]

# Train data is splitted for cross-validation purposes

set.seed(123)

inTrain <- createDataPartition(y = train.df$classe, p = 0.75, list = FALSE)
train <- train.df[inTrain, ]
test <- train.df[-inTrain, ]

```


##Algorithm

A prediction tree has been selected, with a tunning length of 30 iterations. Cross-validation has been done using 25% of the training dataset.

```{r}

fit <- train(classe~., data = train, method = "rpart", tuneLength = 30)
fit

fancyRpartPlot(fit$finalModel)

predicted <- predict(fit, newdata = test)
confusionMatrix(predicted, test$classe)

```

##Predicted values

```{r}

predict(fit, newdata = test.df)

```
